Applied Measure Theory for Probabilistic Modeling

### 2022-05-17 paper/paper.tex
edits

Note the difference from a typical distribution-oriented implementation; we have changed the base measure in order to simplify the log-density. This leads to better compositional efficiency. \chad{"compositional efficiency" is kind of vague, reword}

@@

We could easily have chosen a different base measure and adjusted \verb|logdensity_def| accordingly.
From the perspective of an end user, this could be considered arbitrary.
Therefore, \verb|basemeasure| and \verb|logdensity_def| should be considered implementation details, and not part of the user-facing interface.

@@

Note that a likelihood $\ell$ takes a point in $\Theta$ to an evaluated density in a different space (the observation space $X$). In particular, a likelihood is neither a measure nor a density.

>>>>

Note the difference from a typical distribution-oriented implementation; the log-density consists only of $-x^2/2$, with the normalization term pushed into the base measure.

The normalization term is present for any normal distribution, even those with a different mean or variance. So if we have a product of normals, for example in a regression problem, the normalization can be shared across them. Taking $n$ iid standard \verb|Normal|s as a simple example, this effectively rewrites the log-pdf
\[
    \sum_{j=1}^n \left(-\frac{1}{2}\log 2π -\frac{x_j^2}{2}\right)
    \quad \text{as} \quad
    -\frac{n}{2}\log 2π + \sum_{j=1}^n \left(-\frac{x_j^2}{2}\right)
\]

Further simplification is clearly possible; doing this efficiently in a composable way is the subject of ongoing work.

From our above discussion of moving terms between the log-density and the base measure, the reader may suspect that this can be done arbitrarily, leaving no ``natural'' log-density for a given measure. This is indeed the case; the log-density of a measure is only determined \emph{relative to a particular base measure}.

Our choice of \verb|basemeasure| doesn't reflect any mathematical invariant, only computational convenience. Accordingly, \verb|basemeasure| and \verb|logdensity_def| should be considered implementation details, and not part of the user-facing interface.

@@

@@

Note that a likelihood $\ell$ takes a point in $\Theta$ to an evaluated density in a different space (the observation space $X$). In particular, a likelihood is not a measure.
